---
title: "STAT154_HW2"
author: "Minju Jo"
output: pdf_document
---
###PROBLEM 6###

```{r}
nba<-read.csv("C:/Users/MINJU/Downloads/nba-teams-2019.csv")
head(nba)
```

```{r}
attach(nba)
team<-nba[,'TEAM']
str(nba)
```

```{r}
nba<-nba[,c('W','L','PTS','FGM','X3PM','REB','AST','TOV','STL','BLK')]
str(nba)
```
# we need only 10 columns

##6.1 [Calculation of primary PCA outputs]##

#A#
Obtain the loadings and store them in a matrix, include row and column names.
Display the first five loadings.

```{r}
nba<-scale(nba)
summary(nba)
```
#work with standardized data: variables with mean = 0, and sample variance = 1

```{r}
corNba<-(t(nba)%*%nba)/29
corNba
```

```{r}
loading<-eigen(corNba)$vectors[,1:5]
colnames(loading)<-c('l1','l2','l3','l4','l5')
loading
```
#this is first five loadings

#B#
Obtain the principal components and store them in a matrix, include row and column
names. Display the first five PCs.

```{r}
pc1<-nba%*%loading[,'l1']
rownames(pc1)<-team
pc1
```

```{r}
pc2<-nba%*%loading[,'l2']
rownames(pc2)<-team
pc2
```

```{r}
pc3<-nba%*%loading[,'l3']
rownames(pc3)<-team
pc3
```

```{r}
pc4<-nba%*%loading[,'l4']
rownames(pc4)<-team
pc4
```

```{r}
pc5<-nba%*%loading[,'l5']
rownames(pc5)<-team
pc5
```
#these are 5 pcs

#C#
Obtain the eigenvalues and store them in a vector. Display the entire vector, and
compute their sum.
```{r}
eigenval<-eigen(corNba)$values
eigenval
```

```{r}
sumval<-sum(eigenval)
sumval
```
#this is sum of the eigen values


##6.2 [Eigenvalues]##

#A#
Make a summary table of the eigenvalues: eigenvalue in the first column; proportion
or percentage in the second column; and cumulative percentage in the third column.
Comment on the table: for instance, how do you interpret each eigenvalue?

```{r}
percentage=c()
for (i in 1:length(eigenval)) percentage=c(percentage,(eigenval[i]/sumval)*100)
percentage
```

```{r}
cumulative.percentage=c(percentage[1])
for (i in 2:length(eigenval)) cumulative.percentage=c(cumulative.percentage,cumulative.percentage[i-1]+(eigenval[i]/sumval)*100)
cumulative.percentage
```

```{r}
eigenvalSummary<-data.frame(eigenval, percentage, cumulative.percentage)
rownames(eigenvalSummary)<-paste("eigenval",1:10)
eigenvalSummary
```
#first eigenvalue is the largest and the last one is the smallest
#It is over 90% only with 5 eigenvalues, which can represent almost every data

#B#
Create a scree-plot (with axis labels) of the eigenvalues. What do you see? How do
you read/interpret this chart?


```{r}
plot(percentage,xlab="eigenvalues")
```
#the first eigenvalue explains almost half of the data
#from 6th eigenvalue, it looks less than 5%, which means not that critical


#C#
How do you explain the value of the last eigenvalue? (what¡¯s going on here?)
```{r}
eigenval[10]
```
#it is almost 0

```{r}
percentage[10]
```
#It does not represent the data well

#D#
Based on the different criteria discussed in lecture, how many dimensions (i.e. number
of PCs), would you retain/use? Explain
```{r}
cumulative.percentage
```
I'll keep only five eigenvalues, which is same as 5 dimensions.
Because they already explain 90% of the data which is enough.(By the elbow in scree plot)


##6.3##

#A#
Compute a matrix of correlations between the variables and the PCs. Provide interpretations
for the first four principal components.
```{r}
cor(pc1,nba)
cor(pc2,nba)
cor(pc3,nba)
cor(pc4,nba)
```
From first to third PCs, there is quite high correlation btw the variables and the PCs. However, for the 4th PC, it has less correlation than former ones.

#B#
Plot a Circle of Correlations graph using the first two dimensions (associated to the
first 2 PCs). And provide an interpretation of this graph. Again, recall that you cannot use any PCA-packages.
```{r}
a<-seq(0,2*pi,length.out=100)
plot(cos(a),sin(a),type='l',xlab='dimension1',ylab='dimension2')
arrows(0,0,cor(pc1,nba),cor(pc2,nba),length=0.1,angle=10)
abline(h=0,v=0)
text(cor(pc1,nba),cor(pc2,nba),labels=c('W','L','PTS','FGM','X3PM','REB','AST','TOV','STL','BLK'))
```


##6.4##

#A#
Create a scatter plot of the individuals (i.e. the teams) on the 1st and 2nd PC. Comment on general patterns, as well as on particular patterns.
```{r}
plot(x=pc1,y=pc2, main="PC1 vs PC2", xlab = "PC1", ylab = "PC2")
abline(h = 0, v = 0)
text(pc1,pc2,labels=team)
```

#B#

```{r}
plot(x=pc3,y=pc4, main="PC3 vs PC4", xlab = "PC3", ylab = "PC4")
abline(h = 0, v = 0)
text(pc3,pc4, labels = team)
```

#C#
MIN : negative value in both pc1, pc2 axis. it woudl have some loses
GSW vs NYK : big gap in pc1 axis, almost same in pc2 axis.
gsw won a lot against nyk and got many points.
PHO vs DET : big gap in pc2 axis, almost same in pc1 axis.
DET made a lot of 3 points in game and so there were not many turnovers than PHO.