---
title: "STAT154 HW9"
author: "Minju Jo"
output: pdf_document
---

###PROBLEM 6###

```{r}
library("ISLR")
data=Carseats
```

```{r}
head(data)
```

##a##
```{r}
data$High <- as.numeric(data$Sales>=9)
head(data)
```
create a variable "high"(standard = "sales">=9)

```{r}
data = subset(data, select=-c(ShelveLoc,Urban,US,Sales))
head(data)
```
remove all non-quantative variables + Sales

```{r}
data = scale(data)
```

```{r}
summary(data)
```
data was scaled and centered

```{r}
ndata = nrow(data)
splitIdx = sample(1:ndata, size = nrow(data)*0.7)
train <- as.data.frame(data[splitIdx,])
test <- as.data.frame(data[-splitIdx,])
```

```{r}
dim(train)
dim(test)
```
dataset was split into training/testing set


##b##
```{r}
library(rpart)
```

```{r}
rpTree <- rpart(High~., data=train, method="class", maxdepth=6)
plot(rpTree)
text(rpTree)
```
classification tree

```{r}
err<-mean(predict(rpTree, test, type="class") != test$High)
err
```
evaluate the tree on the train set


##c##
```{r}
da_woHigh <- as.data.frame(data[,-8])
head(da_woHigh)
```
data without high

```{r}
pca <- prcomp(da_woHigh)
pca
```

```{r}
pca_data<-as.data.frame(predict(pca))
pca_data$High <- data[,"High"]
pca_train <- as.data.frame(pca_data[splitIdx,])
pca_test <- as.data.frame(pca_data[-splitIdx,])
```

```{r}
pca_rpTree <- rpart(High~., data=pca_train, method="class", maxdepth=6)
plot(pca_rpTree)
text(pca_rpTree)
```

```{r}
pca_err<-mean(predict(pca_rpTree, pca_test, type="class") != pca_test$High)
pca_err
```
error rate


##d##
```{r}
pca_train3 <- pca_train[,c(1:3,8)]
pca_test3 <- pca_test[,c(1:3,8)]
```

```{r}
head(pca_train3)
head(pca_test3)
```

```{r}
pca_rpTree3 <- rpart(High~., data=pca_train3, method="class", maxdepth=6)
plot(pca_rpTree3)
text(pca_rpTree3)
```

```{r}
pca_err3<-mean(predict(pca_rpTree3, pca_test3, type="class") != pca_test3$High)
pca_err3
```


##e##
```{r}
err
pca_err
pca_err3
```
pca with tree-based methods can reduce err than just using tree-based methods. But we should consider PCs and select how many PCs to use. Because it can decrease or increase misclassification rate.