---
title: "STAT154 HW1"
author: "Minju Jo"
output: pdf_document
---

###PROBLEM 5###
##5A##
```{r}
inner_prod<-function(a,b){
  if(length(a)!=length(b)){
    stop("Length of both vectors should be same!")
  }
  else{
    total=0
    for(i in 1:length(a)){
      total=total+a[i]*b[i]
    }
    return(total)
  }
}
```

```{r}
y=c(1,1,1,1,1)
z=c(2,4,6,8,10)
inner_prod(y,y)
inner_prod(z,z)
inner_prod(y,z)
```

##5B##
```{r}
vnorm<-function(a){
  return(sqrt(inner_prod(a,a)))
}
```

```{r}
b<-c(1,2,3,4,5)
vnorm(b)
```

##5C##
```{r}
unit_norm<-function(x){
  return(x/vnorm(x))
}
```

```{r}
x<-c(1,2,3)
y<-c(1,1,1)
z<-c(1,0,0)
```

```{r}
unit_norm(x)
unit_norm(y)
unit_norm(z)
```

##5D##
```{r}
vector_proj<-function(a,b){
  return((inner_prod(a,b)/inner_prod(b,b))*b)
}
```

```{r}
vector_proj(x,y)
```

##5E##
```{r}
scalar_proj<-function(a,b){
  return(inner_prod(a,b)/vnorm(b))
}
```

```{r}
scalar_proj(x,y)
```


###PROBLEM 6###
##6A##
```{r}
iter=1
mat<-matrix(c(5,-4,3,-14,4,6,11,-4,-3),3,3)
w_old<-c(1,1,1)
s=abs(max(mat%*%w_old))
while(TRUE){
  w_new<-(mat%*%w_old)/s
  s_diff=abs(abs(max(mat%*%w_new))-s)
  s=abs(max(mat%*%w_new))
    
  print(paste('iteration = ',iter))
  print(w_new)
  cat("\n")
  iter=iter+1
    
  w_old<-w_new
  
  if(s_diff<0.001){break}
}
```

```{r}
s
max(eigen(mat)$values)
```
#largest eigenvalue from the Power method and true eigenvalue(from ftn eigen()) are almost same

```{r}
w_old
vnorm(w_old)
```
#eigenvector from the Power method was not scaled. Its norm was 1.12.

```{r}
norm_w_old<-unit_norm(w_old)
norm_w_old
eigen(mat)$vectors[,1]
```
#After normalization(by the ftn unit_norm()), we got the norm_w_old.
#norm_w_old is almost same as (-1)*eigenvector corresponding to eigenvalue 12.

```{r}
ray<-(t(w_old)%*%mat%*%w_old)/t(w_old)%*%w_old
ray
```
#By Rayleigh quotient, we got the almost same value to the largest eigenvalue, which is 12.

```{r}
vnorm(w_old)
```
#It was close to 1, but not 1.

##6B##

```{r}
iter2=1
mat<-matrix(c(5,-4,3,-14,4,6,11,-4,-3),3,3)
w_old2<-c(1,1,1)
s2=vnorm(mat%*%w_old2)

while(TRUE){
  w_new2<-(mat%*%w_old2)/s2
  s_diff2=abs(vnorm(mat%*%w_new2)-s2)
  s2=vnorm(mat%*%w_new2)
    
  print(paste('iteration = ',iter2))
  print(w_new2)
  cat("\n")
  iter2=iter2+1
    
  w_old2<-w_new2
  
  if(s_diff2<0.001){break}
}
```
#I used L2-norm instead of L-infinite norm.
#To calculate L2-norm, I used ftn vnorm().

```{r}
s2
eigen(mat)$values[1]
```
#It is also close to the dominant eigenvalue, 12.

```{r}
w_old2
eigen(mat)$vectors[,1]
```
#I could find that true eigenvector correspoing to eigenvalue 12 and eigenvector through Power test were almost same.

##6C##
```{r}
B<-cbind(c(5,1),c(1,5))
B
```

```{r}
iter3=1
w_old3<-c(1,1)
s3=abs(max(B%*%w_old3))

while(TRUE){
  w_new3<-(B%*%w_old3)/s3
  s_diff3=abs(abs(max(B%*%w_new3))-s3)
  s3=abs(max(B%*%w_new3))
    
  print(paste('iteration = ',iter3))
  print(w_new3)
  cat("\n")
  iter3=iter3+1
    
  w_old3<-w_new3
  
  if(s_diff3<0.001){break}
}
```

```{r}
s3
eigen(B)$values[1]
```
#Dominant eigenvalue from Power Method and true dominant eigenvalue are same.

```{r}
unit_norm(w_new3)
eigen(B)$vectors[,1]
```
#eigenvector is same

```{r}
B1<-B-s3*unit_norm(w_old3)%*%t(unit_norm(w_old3))
B1
```

```{r}
iter4=1
w_old4<-c(2,1)
s4=abs(max(B1%*%w_old4))

while(TRUE){
  w_new4<-(B1%*%w_old4)/s4
  s_diff4=abs(abs(max(B1%*%w_new4))-s4)
  s4=abs(max(B1%*%w_new4))
    
  print(paste('iteration = ',iter4))
  print(w_new4)
  cat("\n")
  iter4=iter4+1
    
  w_old4<-w_new4
  
  if(s_diff4<0.001){break}
}
```

```{r}
s4
eigen(B1)$values[1]
```
#Dominant eigenvalue from Power Method and true dominant eigenvalue are same.

```{r}
unit_norm(w_new4)
eigen(B1)$vectors[,1]
```
#same eigenvectors

```{r}
eigen(B)$values
s3
s4
```
#We get eigenvalues of B through power method with B and B1

```{r}
eigen(B1)$vectors[,1]
unit_norm(w_new3)
unit_norm(w_new4)
```
#We also get eigenvectors of B through power method with B and B1
